---
title: "Take Home Midterm"
author: 
    - Ksenia Gordeeva
    - Rebecca Gordon
    - Havi Khurana
    - Amy Warnock 
    
date: "4/21/2022"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE)


library(tidyverse)
library(rio)
#install.packages("glue")
library(glue)
library(janitor)
#remotes::install_github("datalorax/gapr")
library(gapr)
```

## Part A: Data

### Question 1

Use the function below to download all of the data from the 2014-15 to 2017-18 school years and bind it into a single data frame, using a single function (i.e., one line of code).

```{r a1-download-bind, results = 'hide'}
#function
download_file <- function(year) {
  link <- glue::glue("https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx")
  rio::import(link, setclass = "tibble", na = c("-", "--", "*"))
}

#trying the function out
download_file(15)

#downloading all data and binding into a single data frame
data_raw <- map_dfr(15:18, download_file)
```

Conduct some basic data cleaning to make your data file look like the following.

- Filter for only student groups coded as "White" or "Hispanic/Latino".
- Select variables related to the number of students in each of the levels (1:4), and not percentages or collapsed levels.
- Remove any row that has missing data in any of the n variables

```{r a1-data-cleaning, results = 'hide'}
data <- data_raw %>% 
    clean_names() %>% 
    filter(student_group == "White" | student_group == "Hispanic/Latino") %>% 
    select(academic_year, 
           district, 
           school, 
           student_group, 
           grade_level,
           starts_with("number_level"))  %>% 
    pivot_longer(starts_with("number_level"),
                 names_prefix = "number_level_", 
                 names_to = "level", 
                 values_to = "n") %>% 
    drop_na(n)
```

### Question 2

Sum the *n* across grades and school for each district within each academic year, student group, and level. Calculate the number of schools (unique schools listed) for each district. Also move your data to a wider format where the student groups become new columns, filled by the district *n*. Remove districts that did not report numbers for one or both student groups (i.e., drop missing data from the new columns). Your data frame should look like the below.

```{r a2}
data2 <- data %>%
  group_by(district) %>% 
  mutate(numb_schools = length(unique(school))) %>% 
	count(academic_year, 
	      district, 
	      student_group, 
	      level, 
	      numb_schools, 
	      wt = n) %>%
	pivot_wider(names_from = "student_group",
		          values_from = "n") %>%
	janitor::clean_names() %>%
	drop_na(hispanic_latino, white) %>%
	ungroup() %>% 
  select(academic_year, 
         district, 
         numb_schools, 
         level, 
         everything())

data2
```

## Part B: Achievement Disparities

### Question 1

Estimate the average achievement differences in terms of an effect size for every school district in the state that reported data on both student groups (i.e., using the data we created above), for each academic year. Conduct this analysis twice, once using `group_by() %>% nest() map()` and once using `nest_by() %>% summarize()`.

Note, on my computer these took about 40 seconds per analysis. It will likely take a bit longer for you, but shouldn’t take too long. You may get a warning. Just go ahead and ignore that for now.

```{r b1}

```

### Question 2

Reproduce the plot to confirm that the estimates are equivalent across methods.

```{r b2}

```

### Question 3

Produce a similar plot for each of the first 100 school districts. Once you’ve created all the plots, **reproduce them again using an alternative method**. Make sure you don’t miss the caption noting the number of schools in the district.

Hint: You need to select unique *districts.* You may want to try something like the following:

`dists <- unique(v$district)[1:100]`

`model_list_object %>% filter(district %in% dists)`

```{r b3}

```

#### Extra Credit

For up to 1 point extra credit, create a third set that “staples together” each of the preceding pairs of plots (i.e., to confirm that the plots were produced identically by both methods).
 
```{r b-extra-credit}

```
 
### Question 4

Save the plots into a “plots” directory. Make sure the file names are meaningful.

**HINTS**

- You don’t have to use a loop to create the file names (maybe give `{glue}` a try? Otherwise `paste` or `paste0` will work fine).
- When working with the code, limit the number of plots you’re saving to, say, the first five to make sure it works before running it on all plots.